An Extendable \proglang{Python} Implementation of Robust Optimisation Monte Carlo

### 2021-07-17 latex_files/ch1/1_1_motivation.tex
updated most content in chapter 2,3

\subsubsection*{\textit{Explanation of simulation-based models}}

A simulator-based model is a parameterised stochastic data generating
mechanism \cite{Gutmann2016}. The key characteristic of these
models is that although we can sample (simulate) data points, we
cannot evaluate the likelihood of a specific set of observations
$\data$. Formally, a simulator-based model is described as a
parameterised family of probability density functions
$\{ p_{\yb|\thetab}(\yb) \}_{\thetab}$, whose closed-form is either
unknown or intractable to evaluate. Whereas evaluating
$p_{\yb|\thetab}(\yb)$ is intractable, sampling is
feasible. Practically, a simulator can be understood as a black-box
machine $M_r$\footnote{The subscript $r$ in $M_r$ indicates the
  \textit{random} simulator. In the next chapters we will introduce
  $M_d$ witch stands for the \textit{deterministic} simulator.} that
given a set of parameters $\thetab$, produces samples $\yb$ in a
stochastic manner i.e.\ $M_r(\thetab) \rightarrow \yb$.

Simulator-based models are particularly captivating due to the
modelling freedom they provide; any physical process that can be
conceptualised as a computer program of finite (deterministic or
stochastic) steps can be modelled as a simulator-based model without
any compromise. The modelling freedom includes any amount of hidden
(unobserved) internal variables or logic-based decisions. As always,
this degree of freedom comes at a cost; performing the inference is
particularly demanding from both computational and mathematical
perspective. Unfortunately, the algorithms deployed so far permit the
inference only at low-dimensionality parametric spaces, i.e.\
$\thetab \in \mathbb{R}^D$ where $D$ is small.

\subsubsection*{\textit{Example}}

For underlying the importance of simulator-based models, let us use
the tuberculosis disease spread example as described in
\cite{Tanaka2006}. An overview of the disease spread model is
presented at figure~\ref{fig:tuberculosis_model}. At each stage one of
the following \textit{unobserved} events may happen; (a) the
transmission of a specific haplotype to a new host, (b) the mutation
of an existent haplotype or (c) the exclusion of an infectious host
(recovers/dies) from the population. The random process, which stops
when $m$ infectious hosts are reached\footnote{We suppose that the
  unaffected population is infinite, so a new host can always be added
  until we reach $m$ simultaneous hosts.}, can be parameterised by the
transmission rate $\alpha$, the mutation rate $\tau$ and the exclusion
rate $\delta$, creating a $3D$-parametric space
$\thetab = (\alpha, \tau, \delta)$. The outcome of the process is a
variable-size tuple $\yb_{\thetab}$, containing the population
contaminated by each different haplotype, as described in
figure~\ref{fig:tuberculosis_model}. Let's say that the disease has
been spread in a real population and when $m$ hosts were contaminated
simultaneously, the vector with the infectious populations has been
measured to be $\data$. We would like to discover the parameters
$\thetab = (\alpha, \tau, \delta)$ that generated the spreading
process and led to the specific outcome $\data$. Computing
$p(\yb=\data|\thetab)$ requires tracking all tree-paths that could
generate the specific tuple; such exhaustive enumeration becomes
intractable when $m$ grows larger, as in real-case scenarios. In
figure~\ref{fig:tuberculosis_model} we can observe that a transmission
followed by a recovery/death creates a loop, reinstating the process
to the previous step, which also complicates the exhaustive
enumeration. Hence, representing the process with a simulator-based
model\footnote{which is simple and efficient} and performing
likelihood-free inference is the recommended solution.

\begin{figure}[!ht]
    \begin{center}
      \includegraphics[width=0.75\textwidth]{./latex_files/images/chapter1/tuber_model_1.png}
    \end{center}
    \caption[The tuberculosis spreading process.]{Depiction of a random example from the tuberculosis
      spreading process. The image has been taken from
      \cite{Lintusaari2017}.}
    \label{fig:tuberculosis_model}
\end{figure}

\subsubsection*{\textit{Goal of Simulation-Based Models}}

As in most Machine Learning (ML) concepts, the fundamental goal is the
derivation of one(many) parameter configuration(s) $\thetab^*$ that
\textit{describe} the data best i.e.\ generate samples
$M_r(\thetab^*)$ that are as close as possible to the observed data
$\data$. In our case, following the approach of Bayesian ML, we treat
the parameters of interest $\thetab$ as random variables and we try to
\textit{infer} a posterior distribution $p(\thetab|\data)$ on them.

\subsubsection*{\textit{Robust Optimisation Monte Carlo (ROMC) method}}

The ROMC method~\cite{Ikonomov2019} is very a recent
likelihood-free approach. Its fundamental idea is the transformation
of the stochastic data generation process $M_r(\thetab)$ to a
deterministic mapping $g_i(\thetab)$, by sampling the variables that
produce the randomness $\vb_i \sim p(\V)$. Formally, in every
stochastic process the randomness is influenced by a vector of random
variables $\V$, whose state is unknown prior to the execution of the
simulation; sampling the state makes the procedure deterministic,
namely $g_i(\thetab) = M_d(\thetab, \V=\vb_i)$. This approach
initially introduced at \cite{Meeds2015} with the title
\textit{Optimisation Monte Carlo (OMC)}. The ROMC extended this
approach by resolving a fundamental failure-mode of OMC\@. The ROMC
describes a methodology for approximating the posterior through a
series of steps, without explicitly enforcing which algorithms must be
utilised for each step\footnote{The implementation chooses a specific
  algorithm for each task, but this choice has just a demonstrative
  value; any appropriate algorithm can be used instead.}; in this
sense, it can be perceived as a meta-algorithm.

\subsubsection*{\textit{Implementation}}

The most important contribution of this work is the implementation of
the ROMC method in the Python package Engine for Likelihood-Free
Inference (ELFI) \cite{1708.00707}. Since the method has been
published quite recently, it has not been implemented by now in any ML
software. This work attempts to provide the research community with a
robust and extensible implementation for further experimentation.

>>>>

An implicit or simulator-based model is a parameterised stochastic
data generating mechanism \cite{Gutmann2016}. The key characteristic
of these models is that although we can sample (simulate) data points,
we cannot evaluate the likelihood of a specific set of observations
$\data$. Formally, a simulator-based model is described as a
parameterised family of probability density functions
$\{ p_{\yb|\thetab}(\yb) \}_{\thetab}$, whose closed-form is either
unknown or intractable to evaluate. Whereas evaluating
$p_{\yb|\thetab}(\yb)$ is intractable, sampling is
feasible. Practically, a simulator can be understood as a black-box
machine $M_r$\footnote{The subscript $r$ in $M_r$ indicates the
  \textit{random} simulator. In the next chapters we will introduce
  $M_d$ indicating the \textit{deterministic} simulator.} that
given a set of parameters $\thetab$, produces samples $\yb$ in a
stochastic manner i.e.\ $M_r(\thetab) \rightarrow \yb$.

Simulator-based models are particularly captivating due to the
modelling freedom they provide; any physical process that can be
conceptualised as a computer program of finite steps can be modelled
as a simulator-based model without any compromise. The modelling
freedom includes any amount of hidden (unobserved) internal variables
or rule-based decisions. Hence, implicit models are often used to
model physical phenomena in many fiels such as biology, neuroscience
etc. This degree of freedom comes at a cost; performing the inference
is particularly demanding from both the computational and the
mathematical perspective. The class of methods that try to approximate
the posterior distribution in these cases are called likelihood-free
inference methods. Unfortunately, the algorithms deployed so far
permit the inference only at low-dimensional parametric spaces,
i.e.\ $\thetab \in \mathbb{R}^D$ where $D$ is small\footnote{In broad terms, small refers to dimensionality $D$ less than a hundred.}.

\subsubsection*{\textit{Goal of implicit models}}

As in most Machine Learning (ML) concepts, the fundamental goal is the
derivation of one (or many) parameter configuration(s) $\thetab^*$
that \textit{describe} the data best i.e.\ generate
samples $M_r(\thetab^*)$ that are as close as possible to the observed
data $\data$. In our case, following the Bayesian approach, we treat
the parameters of interest $\thetab$ as random variables and we try to
\textit{infer} a posterior distribution $p(\thetab|\data)$ on them.

### 2021-07-17 latex_files/ch3/3_2_1_general_design.tex
updated most content in chapter 2,3

In figure \ref{fig:romc_overview} we present an overview of our
implementation; one may interpret figure~\ref{fig:romc_overview} as a
depiction of the main class of our implementation, called
\pinline{ROMC}, while the entities inside the green and blue
ellipses are the main functions of the class. Following Python's
naming principles, the methods starting with an underscore (green
ellipses) represent internal (private) functions and are not meant to
be used by a user, whereas the rest of the methods (blue ellipses) are
the functionalities the user interacts with. As mentioned before, the
implementation favours extensibility; the building blocks that compose
the method have been designed in a modular fashion so that a
practitioner may replace them without the method to collapse.

Figure \ref{fig:romc_overview} groups the ROMC implementation into the
training, the inference and the evaluation part. The training part includes all the
steps until the computation of the proposal regions; sampling the
nuisance variables, defining the optimisation problems, solving them,
constructing the regions and fitting local surrogate models. The
inference part comprises of evaluating the unnormalised posterior (and
the normalised one, in low-dimensional cases), sampling and computing
an expectation. Moreover, the ROMC implementation provides some
utilities for inspecting the training process, such as plotting the
histogram of the distances
$d^*_i = g_i(\theta_i^*), \: \forall i \in \{1, \ldots, n_1 \}$ after
solving the optimisation problems and visualising the constructed
bounding box\footnote{if the parametric space is up to $2D$}. Finally,
two functionalities for evaluating the inference are implemented; (a)
computing the Effective Sample Size (ESS) of the weighted samples and
(b) measuring the divergence between the approximate posterior the
ground-truth, if the latter is available.\footnote{Normally, the
  ground-truth posterior is not available; However, this functionality
  is useful in cases where the posterior can be computed numerically
  or with an alternative method (i.e.\ ABC Rejection Sampling) and we
  would like to measure the discrepancy between the two
  approximations.}

\subsubsection*{Parallelising the processes}

As stated above, the most critical advantage of the ROMC method is that it can be fully parallelised. In our implementation, we have exploited this fact by implementing a parallel version in the following tasks; (a) solving the optimisation problems, (b) constructing bounding box regions, (c) sampling and (d) evaluating the posterior. Parallelism has been achieved using the package \pinline{multiprocessing}. The specific package enables concurrency, using subprocesses instead of threads, hence side-stepping the Global Interpreter (GIL). In our implementation we use the \pinline{Pool} object, which offers a convenient means of parallelising the execution of a function across multiple input values (data parallelism). For activating the parallel version of the algorithms, the user has to just pass the argument \pinline{parallelize=True} at the initialisation of the \pinline{ROMC} method.

@@

For illustrating the functionalities we choose as running example the
following model, introduced by~\cite{Ikonomov2019},

\begin{gather} \label{eq:1D_example} p(\theta) =
\mathcal{U}(\theta;-2.5,2.5)\\ p(y|\theta) = \left\{
    \begin{array}{ll} \theta^4 + u & \mbox{if } \theta \in [-0.5, 0.5]
\\ |\theta| - c + u & \mbox{otherwise}
    \end{array} \right.\\ u \sim \mathcal{N}(0,1)
\end{gather}

@@

In the model \eqref{eq:1D_example}, the prior is the uniform
distribution in the range $[-2.5, 2.5]$ and the likelihood a Gaussian
distribution. There is only one observation $y_0 = 0$. The inference
in this particular example can be performed quite easily, without
incorporating a likelihood-free inference approach. We can exploit this
fact for validating the accuracy of our implementation. The
ground-truth posterior, approximated computationally, is shown in
figure \ref{fig:example_gt}.

\begin{figure}[h]
    \begin{center}
\includegraphics[width=0.7\textwidth]{./latex_files/images/chapter3/example_gt.png}
    \end{center}
  \caption[Ground-truth posterior distribution of the simple 1D
example.]{Ground-truth posterior distribution of the simple 1D
example.}
  \label{fig:example_gt}
\end{figure}

\subsubsection*{ELFI code for modelling the example}

In the following code snippet, we code the model at \textit{ELFI} and
we initialise the ROMC inference method. We observe that the
initialisation of the ROMC inference method is quite intuitive; we
just pass the final (distance) node of the simulator as argument, as
in all $\textit{ELFI}$ inference methods. The argument
\pinline{bounds}, although optional, is important for many
functionalities (e.g. approximating the partition function, setting
the bounds of the Bayesian optimisation etc.) so it is recommended to
be passed.

\begin{pythoncode}
  import elfi import scipy.stats as ss
  import numpy as np

@@

  # Initialise the ROMC inference method
  bounds = [(-2.5, 2.5)] # limits of the prior
  parallelize = True # activate parallel execution
  romc = elfi.ROMC(d, bounds=bounds, parallelize=parallelize)
\end{pythoncode}

\begin{figure}[!ht]
    \begin{center}
\includegraphics[width=0.8\textwidth]{./latex_files/graphs/ROMC.png}
    \end{center}
    \caption[Overview of the ROMC implementation.]{Overview of the
ROMC implementation. The training part follows a sequential pattern;
the functions in the green ellipses must be called in a sequential
fashion for completing the training part and define the posterior
distribution. The functions in blue ellipses are the functionalities
provided to the user.}
    \label{fig:romc_overview}
\end{figure}

>>>>

The overview of our implementation is illustrated in
figure~\ref{fig:romc_overview}; one may interpret
figure~\ref{fig:romc_overview} as a depiction of the main class and
the ellipses are the main routines of the class. Following
\proglang{Python}'s naming principles, the methods starting with an
underscore (green ellipses) represent internal (private) functions,
whereas the rest (blue ellipses) are the public methods that the user
interacts with.

Figure~\ref{fig:romc_overview} splits ROMC into the training, the
inference and the evaluation part. The training part includes all the
steps until the computation of the proposal regions i.e. sampling the
nuisance variables, defining the optimization problems, solving them,
constructing the regions and fitting local surrogate models. The
inference part comprises of evaluating the unnormalized posterior (and
the normalized when is possible), sampling and computing an
expectation. From an algorithmic point of view, the training and the
inference parts are \emph{isolated}. We also provide some utilities
for inspecting the training process, such as plotting the histogram of
the final distances or visualizing the constructed bounding
boxes. Finally, in the evaluation part, we provide two methods for
evaluating the inference; (a) computing the Effective Sample Size
(ESS) of the samples and (b) measuring the divergence between the
approximate posterior the ground-truth, if the latter is
available.\footnote{Normally, the ground-truth posterior is not
  available; However, this functionality is useful in cases where the
  posterior can be computed numerically or with an alternative method
  (i.e.\ ABC Rejection Sampling), and we would like to measure the
  discrepancy between the two approximations.}

\begin{figure}[H]
    \begin{center}
\includegraphics[width=0.8\textwidth]{./latex_files/graphs/ROMC.png}
    \end{center}
    \caption[Overview of the ROMC implementation.]{Overview of the
ROMC implementation. The training part follows a sequential pattern;
the functions in the green ellipses must be called in a sequential
fashion for completing the training part and define the posterior
distribution. The functions in blue ellipses are the functionalities
provided to the user.}
    \label{fig:romc_overview}
\end{figure}

\subsubsection*{Parallelizing ROMC method}

ROMC has the significant advantage of being fully parallelizable. We
exploit this fact by implementing a parallel version of all the major
components of the method; (a) solving the optimization problems, (b)
constructing bounding box regions, (c) sampling and (d) evaluating the
posterior. We parallelize these processes using the built-in
\proglang{Python} package \pkg{multiprocessing}. The specific package
enables concurrency, using subprocesses instead of threads, for
side-stepping the Global Interpreter (GIL). For activating the
parallel version of the algorithm, the user simpy has to pass the
argument \code{parallelize=True} when instantiating \code{ROMC}.

\begin{Code}
---------------------------------- python ----------------------------------
>>> elfi.ROMC(<output_node>, parallelize=True)
----------------------------------------------------------------------------
\end{Code}

@@

For illustrating the functionalities we will use the following running
example introduced by~\cite{Ikonomov2019},

\begin{gather} \label{eq:1D_example}
  p(\theta) = \mathcal{U}(\theta;-2.5,2.5)\\ \label{eq:1D_example_eq_2}
  p(y|\theta) = \left\{
    \begin{array}{ll} \theta^4 + u & \mbox{if } \theta \in [-0.5, 0.5]
\\ |\theta| - c + u & \mbox{otherwise}
    \end{array} \right.\\ 
  u \sim \mathcal{N}(0,1)
\end{gather}

@@

The prior is a uniform distribution in the range $[-2.5, 2.5]$ and the
likelihood is defined at~\ref{eq:1D_example_eq_2}. There is only one
observation $y_0 = 0$. The inference in this particular example can be
performed quite easily without using a likelihood-free inference
approach. We can exploit this fact for validating the accuracy of our
implementation. The ground-truth posterior approximated
computationally is shown in figure~\ref{fig:example_gt}.

\begin{figure}[!h]
    \begin{center}
\includegraphics[width=0.7\textwidth]{./latex_files/images/chapter3/example_gt.png}
    \end{center}
  \caption[Ground-truth posterior distribution of the simple 1D
example.]{Ground-truth posterior distribution of the simple 1D
example.}
  \label{fig:example_gt}
\end{figure}

In the following code snippet, we code the model at \pkg{ELFI}.

\begin{Code}
------------------------------ python snippet ------------------------------
  import elfi import scipy.stats as ss
  import numpy as np

@@

  # Initialise the ROMC inference method
  bounds = [(-2.5, 2.5)] # limits of the prior
  parallelize = True # activate parallel execution
  romc = elfi.ROMC(d, bounds=bounds, parallelize=parallelize)
----------------------------------------------------------------------------    
\end{Code}

### 2021-07-17 latex_files/ch3/3_2_2_training.tex
updated most content in chapter 2,3

The training part contains the following 6 functionalities:

\begin{enumerate}
\item \mintinline{python}{romc.solve_problems(n1, use_bo=False, optimizer_args=None, seed=None)}
\item \mintinline{python}{romc.estimate_regions(eps_filter,}

      \mintinline{python}{                      use_surrogate=None, region_args=None,}

      \mintinline{python}{                      fit_models=False, fit_models_args=None,}

      \mintinline{python}{                      eps_region=None, eps_cutoff=None)}

    \item \mintinline{python}{romc.fit_posterior(n1, eps_filter, use_bo=False, optimizer_args=None,}

          \mintinline{python}{                   seed=None, use_surrogate=None, region_args=None,}

          \mintinline{python}{                   fit_models=False, fit_models_args=None,}

          \mintinline{python}{                   eps_region=None, eps_cutoff=None)}

\item \mintinline{python}{romc.distance_hist(savefig=False, **kwargs)}
\item \mintinline{python}{romc.visualize_region(i, savefig=False)}
\item \mintinline{python}{romc.compute_eps(quantile)}
\end{enumerate}

\subsubsection*{Function (i): Define and solve the optimisation problems}

\pinline{romc.solve_problems(n1, use_bo=False, optimizer_args=None, seed=None)}
\vspace{5mm}

\noindent
This routine is responsible for (a) drawing the nuisance variables,
(b) defining the optimisation problems and (c) solving them using either a
gradient-based optimiser or Bayesian optimisation. The aforementioned
tasks are done in a sequential fashion, as show in
figure~\ref{fig:romc_overview}. The definition of the optimisation
problems is performed by drawing $n_1$ integer numbers from a discrete
uniform distribution $u_i \sim \mathcal{U}\{1, 2^{32}-1\}$. Each
integer $u_i$ is the seed used in ELFI's random simulator. Hence from
an algorithmic point-of-view drawing the state of all random
variables $\vb_i$ as described in the previous chapter, traces back to
just setting the seed that initialises the state of the pseudo-random
generator, before asking a sample from the simulator.

Finally, passing an integer number as the argument \pinline{seed}
absorbs all the randomness of the optimisation part (e.g.\ drawing
initial points for the optimisation procedure), making the whole
process reproducible and deterministic.

Setting the argument \pinline{use_bo=True}, chooses the Bayesian
Optimisation scheme for obtaining $\thetab_i^*$. In this case, apart
from obtaining the optimal points $\thetab_i^*$, we also fit a Gaussian
Process (GP) as surrogate model $\hat{d}_i$. In the following steps,
$\hat{d}_i(\thetab)$ will replace $g_i(\thetab)$ when calling the
indicator function.

\subsubsection*{Function (ii): Construct bounding boxes and fit local surrogate models}

\mintinline{python}{romc.estimate_regions(eps_filter,}

      \mintinline{python}{                      use_surrogate=None, region_args=None,}

      \mintinline{python}{                      fit_models=False, fit_models_args=None,}

      \mintinline{python}{                      eps_region=None, eps_cutoff=None)}
\vspace{5mm}

This routine constructs the bounding boxes around the optimal points
$\thetab_i^* : i = 1, 2, \ldots, n_1$ following
Algorithm~\ref{alg:region_construction}. The Hessian matrix is
approximated based on the Jacobian $\hess_i = \jac_i^T \jac_i$. The
eigenvectores are computed using the function
\pinline{numpy.linalg.eig()} that calls, under the hood, the
\pinline{_geev LAPACK}. A check is performed so that the matrix
$\hess_i$ is not singular; if this is the case, the eigenvectors are
set to be the vectors of the standard Euclidean basis i.e.\
$\{ \mathbf{e_1} = (1, 0, \ldots), \mathbf{e_2} = (0,1,0,\ldots),
\text{etc} \}$. Afterwards, the limits are obtained by repeteadly querying
the distance function ($g_i(\thetab)$ or $\hat{d}(\thetab)$) along the
search directions. In section \ref{subsec:developers}, we provide some
details regarding the way the bounding box is defined as a class and
sampling is performed on it.

\subsubsection*{Function (iii): Perform all training steps in a single call}

\mintinline{python}{romc.fit_posterior(n1, eps_filter, use_bo=False, optimizer_args=None,}

          \mintinline{python}{                seed=None, use_surrogate=None, region_args=None,}

          \mintinline{python}{                fit_models=False, fit_models_args=None,}

          \mintinline{python}{                eps_region=None, eps_cutoff=None)}

\vspace{5mm}
\noindent

This function merges all steps for constructing the bounding box into
a single command. If the user doesn't want to manually inspect the
histogram of the distances before deciding where to set the threshold
$\epsilon$, he may call \pinline{romc.fit_posterior()} and the whole
training part will be done end-to-end. There are two alternatives for
setting the threshold $\epsilon$; the first is to set to a specific
value blindly and the second is to set at as a specific quantile of
the histogram of distances. In the second scenario the
\pinline{quantile} argument must be set to a floating number in the
range $[0,1]$ and \pinline{eps='auto'}.

\subsubsection*{Function (iv): Plot the histogramm of the optimal points}

\pinline{romc.distance_hist(**kwargs)}
\vspace{5mm}
\noindent

This function can serve as an intermediate step of manual inspection,
for helping the user choose which threshold $\epsilon$ to use. It
plots a histogram of the distances at the optimal point
$g_i(\thetab_i^*) : \{i = 1, 2, \ldots, n_1\}$ or
$d_i^*$ in case \pinline{use_bo=True}. The function accepts all
keyword arguments and forwards them to the underlying
\pinline{matplotlib.hist()} function; in this way the user may
customise some properties of the histogram, such as the number of bins
or the range of values.

\subsubsection*{Function (v): Plot the acceptance region of the objective functions}

\pinline{romc.visualize_region(i)}
\vspace{5mm}
\noindent

It can be used as an inspection utility for cases where the parametric
space is up to two dimensional. The argument $\pinline{i}$ is the
index of the corresponding optimization problem i.e.\ $i<n_1$.

\subsubsection*{Function (vi): Compute $\epsilon$ automatically based on the distribution of $d^*$}

\pinline{romc.compute_eps(quantile)}
\vspace{5mm}
\noindent

\noindent
This function return the appropriate distance value $d_{i=\kappa}^*$
where $\kappa = \lfloor \frac{quantile}{n} \rfloor$ from the
collection $\{ d_i^* \} \forall i = \{1, \ldots, n\}$ where $n$ is the
number of accepted solutions. It can be used to automate the selection
of the threshold $\epsilon$, e.g.\
\pinline{eps=romc.compute_eps(quantile=0.9)}.

@@

Here we will illustrate the aforementioned functionalities using the
simple 1D example we set up in the previous section. The following
code snippet performs the training part in ELFI.

\begin{pythoncode}
  n1 = 500 # number of optimisation problems
  seed = 21 # seed for solving the optimisation problems
  eps = .75 # threshold for bounding box
  use_bo = False # set to True for switching to Bayesian optimisation

  # Training step-by-step
  romc.solve_problems(n1=n1, seed=seed, use_bo=use_bo)
  romc.theta_hist(bins=100)
  romc.estimate_regions(eps=eps)
  romc.visualize_region(i=1)

  # Equivalent one-line command
  # romc.fit_posterior(n1=n1, eps=eps, use_bo=use_bo, seed=seed)
\end{pythoncode}

As stated before, switching to the Bayesian optimisation scheme needs
nothing more the setting the argument \pinline{use_bo=True}; all the
following command remain unchanged. In figure
\ref{fig:example_training_hist} we illustrate the distribution of the
distances obtained and the acceptance area of the first optimisation
problem. We observe that most optimal points produce almost zero
distance.

>>>>

The training part contains six methods.

\begin{Code}
---------------------------------- python ----------------------------------
>>> romc.solve_problems(n1,
                        use_bo=False,
                        optimizer_args=None,
                        seed=None)
----------------------------------------------------------------------------                    
\end{Code}

\noindent
This method (a) draws the nuisance variables, (b) defines the
optimisation problems and (c) solves them using either a
gradient-based optimizer or Bayesian optimization. These tasks are
compleed sequentially, as shown in figure~\ref{fig:romc_overview}. The
definition of the optimisation problems is performed by drawing
\code{n_1} integer numbers from a discrete uniform distribution
$u_i \sim \mathcal{U}\{1, 2^{32}-1\}$. Each integer $u_i$ is the seed
used in \pkg{ELFI}'s random simulator. The seed $u_i$ is the
computational analogous of the random variables $\vb_i$, that were
defined in the previous chapter; freezing the \code{seed} absorbs all
the randomness of the simulator transforming it to a deterministic
function. Setting the argument \code{use_bo=True}, chooses the
Bayesian Optimisation scheme for obtaining $\thetab_i^*$. In this
case, apart from obtaining the optimal points $\thetab_i^*$, we also
fit a Gaussian Process (GP) as surrogate model $\hat{d}_i$. In this
scenario, in the following steps, $\hat{d}_i(\thetab)$ will replace
$g_i(\thetab)$ as the distance function.

\begin{Code}
---------------------------------- python ----------------------------------
>>> romc.distance_hist(**kwargs)
----------------------------------------------------------------------------  
\end{Code}

\noindent
This function helps the user decide which threshold $\epsilon$ to
use. It plots a histogram of the distances at the optimal point
$g_i(\thetab_i^*) : \{i = 1, 2, \ldots, n_1\}$ or $d_i^*$ in case
\code{use_bo=True}. The function accepts all keyword arguments and
forwards them to the underlying \code{matplotlib.hist()} function; in
this way the user may customise some properties of the histogram, such
as the number of bins or the range of values.

\begin{Code}
---------------------------------- python ----------------------------------  
>>> romc.estimate_regions(eps_filter,
                          use_surrogate=None,
                          region_args=None,
                          fit_models=False,
                          fit_models_args=None,
                          eps_region=None,
                          eps_cutoff=None)
----------------------------------------------------------------------------
\end{Code}

\noindent
This method constructs a bounding box around the optimal point
$\thetab_i^* : i = 1, 2, \ldots, n_1$ following
Algorithm~\ref{alg:region_construction}. The Hessian matrix is
approximated based on the Jacobian $\hess_i = \jac_i^T \jac_i$. The
eigenvectores are computed using the function
\code{numpy.linalg.eig()}. A check is performed so that the matrix
$\hess_i$ is not singular; if this is the case, the eigenvectors are
set to be the vectors of the standard Euclidean basis i.e.\
$\{ \mathbf{e_1} = (1, 0, \ldots), \mathbf{e_2} = (0,1,0,\ldots),
\text{etc} \}$. The limits of the bounding box are obtained by
repeteadly querying the distance function ($g_i(\thetab)$ or
$\hat{d}(\thetab)$) along the search directions given by $\hess_i$.

\begin{Code}
---------------------------------- python ----------------------------------
>>> romc.fit_posterior(n1,
                       eps_filter,
                       use_bo=False,
                       optimizer_args=None,
                       seed=None,
                       use_surrogate=None,
                       region_args=None,
                       fit_models=False,
                       fit_models_args=None,
                       eps_region=None,
                       eps_cutoff=None)
----------------------------------------------------------------------------
\end{Code}

\noindent
The function \code{fit_posterior} is a syntactic sugar for applying
\code{solve_problems} and \code{estimate_regions} into a single
step.

% merges all steps for constructing the bounding box into
% a single command. If the user doesn't want to manually inspect the
% histogram of the distances before deciding where to set the threshold
% $\epsilon$, he may call \code{romc.fit_posterior()} and the whole
% training part will be done end-to-end. There are two alternatives for
% setting the threshold $\epsilon$; the first is to set to a specific
% value blindly, and the second is to set at as a specific quantile of
% the histogram of distances. In the second scenario the
% \code{quantile} argument must be set to a floating number in the
% range $[0,1]$ and \code{eps='auto'}.

\begin{Code}
---------------------------------- python ----------------------------------
>>> romc.visualize_region(i)
----------------------------------------------------------------------------
\end{Code}

\noindent

It can be used for plotting the bounding box around the optimal point,
when the parameter space is $1D$ or $2D$. The argument \code{i} is the
index of the corresponding optimization problem i.e.\ $i<n_1$.

\begin{Code}
---------------------------------- python ----------------------------------  
>>> romc.compute_eps(quantile)
----------------------------------------------------------------------------
\end{Code}

\noindent
This function return the appropriate distance value $d_{i=\kappa}^*$
where $\kappa = \lfloor \frac{quantile}{n} \rfloor$ from the
collection $\{ d_i^* \} \forall i = \{1, \ldots, n\}$ where $n$ is the
number of accepted solutions. It can be used to automate the selection
of the threshold $\epsilon$, e.g.\
\code{eps=romc.compute_eps(quantile=0.9)}.

@@

In the following snippet, we put together the routines described above
to perform the training part at our running example.

\begin{Code}
------------------------------ python snippet ------------------------------  
  n1 = 500 # number of optimisation problems
  seed = 21 # seed for solving the optimisation problems
  eps = .75 
  use_bo = False # set to True for switching to Bayesian optimisation

  # Training step-by-step
  romc.solve_problems(n1=n1, seed=seed, use_bo=use_bo)
  romc.theta_hist(bins=100) # plot hist to decide which eps to use

  eps = .75 # threshold for the bounding box based on histogram inspection
  romc.estimate_regions(eps=eps) # build the bounding boxes

  romc.visualize_region(i=1) # for inspecting visually the bounding box

  # Equivalent one-line command
  # romc.fit_posterior(n1=n1, eps=eps, use_bo=use_bo, seed=seed)
----------------------------------------------------------------------------  
\end{Code}

In figure \ref{fig:example_training_hist} we illustrate the
distribution of the distances obtained and the acceptance area of the
first optimisation problem. We observe that most optimal points
produce almost zero distance.
